\chapter{Conclusion}
\label{conclusion}

We designed a model of the Linux page cache and implemented it in the
\simgrid-based \wrench simulation framework to simulate the execution
of distributed applications.
Evaluation results show that our model improves simulation accuracy
substantially, reducing absolute relative simulation errors by up to
9$\times$ (see results of the single-threaded experiment). The
availability of asymmetrical disk bandwidths in the forthcoming
\simgrid release will further improve these results.
Our page cache model is publicly available in the \wrench GitHub
repository.

Page cache simulation can be instrumental in a number of studies. For
instance, it is now common for HPC clusters to run applications in
Linux control groups (cgroups), where resource consumption is limited,
including memory and therefore page cache usage. Using our simulator,
it would be possible to study the interaction between memory allocation
and I/O performance, for instance to improve scheduling algorithms or
avoid page cache starvation~\cite{zhuang2017}. Our simulator could also
be leveraged to evaluate solutions that reduce the impact of network
file transfers on distributed applications, such as burst
buffers~\cite{ferreiradasilva-fgcs-bb-2019}, hierarchical file
systems~\cite{islam2015triple}, active storage~\cite{5496981}, or
specific hardware architectures~\cite{hayot2020performance}. 

Not all I/O behaviors are captured by currently available simulation models,
including the one developed in this work, 
which could substantially limit the accuracy of simulations.
Relevant extensions to this work include more
accurate descriptions of anonymous memory usage in applications, 
which strongly affects I/O times through writeback cache. File access patterns
 might also be worth including in the simulation models,
as they directly affect page cache content.
% readahead and persistent storage could also be adeed
% Simulation results could be made even more accurate by a deeper
% investigation of page cache flushing and eviction order. 

All the results of this work have been collected and made available at the Github repository 
\url{https://github.com/big-data-lab-team/paper-io-simulation} 
with scripts and Jupyter notebooks to generate and view figures. 
Chapter~\ref{method}, Chapter~\ref{experiment} in this thesis have been 
submitted to as a pre-print to arXiv and available at: 
\url{https://arxiv.org/abs/2101.01335}. 

\subsubsection{Contribution of authors}

\noindent Hoang-Dung Do: performing experiments; analyzing the results; building, 
implementing and integrating the page cache simulation model in \wrench; 
writing, editing, polishing of the paper.

\noindent Val\'erie Hayot-Sasson: providing background knowledge and documents; 
guiding experiments; analyzing the results; polishing the paper.

\noindent Rafael Ferreira da Silva: consulting, reviewing the implementation of the model 
in \wrench.

\noindent Christopher Steele: providing real-world use cases and nighres workflows.

\noindent Henri Casanova: consulting, reviewing, refining the implementation of the 
model in \wrench ; performing unit tests; polishing, and proofing of the paper.

\noindent Tristan Glatard: research supervisor; funding; refining the initial ideas; 
guiding experiments; editing, writing, polishing, and proofing of the paper.